---
title: "Mapping Assignment"
author: 'Team 7: Rong Li, Minqi Li, Zhiwei Liang'
date: "10/28/2020"
output: pdf_document
---

```{r read_data, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(warning = FALSE)
library(tidyverse)
library(knitr)    
library(stringr)
library(ggmap)
library(magrittr)
library(dplyr)
library(drat)
library(hurricaneexposuredata)
library(hurricaneexposure)
library(magrittr)
library(tmap)
library(tmaptools)
library(maps)
library(usmap)
library(sp)
library(sf)
library(leaflet)
addRepo("geanders")
```


## The first map for Floyd-1999 made with ggplot2

```{r, warning=FALSE}
## load data
data("hurr_tracks")
data("rain")
data(county.fips)

MainStates <- map_data("state")
AllCounty <- map_data("county")
counties34 <- c("alabama", "arkansas", "connecticut", "delaware", "district of columbia",
                "florida","georgia", "illinois", "indiana","iowa", "kansas", "kentucky", 
                "louisiana","maine", "maryland","massachusetts", "michigan", "mississippi",
                "missouri","new hampshire", "new jersey", "new york", "north carolina", 
                "ohio", "oklahoma", "pennsylvania", "rhode island", "south carolina",
                "tennessee", "texas","vermont","virginia", "west virginia", "wisconsin")
my_states <- subset(MainStates, region %in% counties34)
my_counties <- subset(AllCounty, region %in% counties34)

## get the track data of storm Floyd-1999
Floyd <- hurr_tracks[which(hurr_tracks$storm_id == "Floyd-1999"),]

## get the rain data of storm Floyd-1999
Floyd_rain <- rain %>% filter(storm_id == "Floyd-1999")
sum_precip <- Floyd_rain %>% group_by(fips) %>% dplyr::summarise(Rainfall = sum(precip))
rain_county <- fips_info(sum_precip$fips)
Floyd_rain <- merge(sum_precip, rain_county, by = 'fips')
Floyd_rain$county %<>% str_replace(" County","")
Floyd_rain$full %<>% tolower()
Floyd_rain$county %<>% tolower()
Floyd_rain %<>% dplyr::rename(region=full, subregion=county)
a <- subset(Floyd_rain, region %in% counties34)
rainmap <- left_join(a, my_counties, by = c("region", "subregion"))

rainmap <- rainmap %>% mutate('Rainfall (mm)' = cut(rainmap$Rainfall, 
                                                   breaks = c(0, 25, 50, 75, 100, 125, 150, 175, 200, 225), 
                                                   include.lowest = TRUE))

ggplot() + 
  geom_polygon(data=my_counties, aes(x=long, y=lat, group=group),
               color="gray", fill="white", size = .1 ) + 
  geom_polygon(data = rainmap, aes(x = long, y = lat, group = group, 
                                   fill = `Rainfall (mm)`), 
               color = "grey", size = 0.2, alpha = 1.6) + 
  geom_polygon(data=my_states, aes(x=long, y=lat, group=group),
               color="black", fill="white",  size = .5, alpha = .3) + 
  geom_path(aes(x = Floyd$longitude, y = Floyd$latitude), color = "red", size=0.4) + 
  scale_fill_brewer(palette = "Blues") + 
  xlim(-108, -65) + 
  ylim(23, 50) + 
  ggtitle("Floyd-1999") +
  theme(plot.title = element_text(hjust = 0.5))
```


## The second map for Allison-2001 made with ggplot2

```{r,warning=FALSE}
## get the track data of storm Allison-2001
Allison <- hurr_tracks[which(hurr_tracks$storm_id == "Allison-2001"),]

## get the rain data of storm Floyd-1999
Allison_rain <- rain %>% filter(storm_id == "Allison-2001")
sum_precip <- Allison_rain %>% group_by(fips) %>% dplyr::summarise(Rainfall = sum(precip))
sum_precip$fips %<>% as.numeric()
Allison_rain <- left_join(sum_precip, county.fips, by = 'fips')
Allison_rain %<>% separate(polyname, c("region", "subregion"), sep=",")
rainmap <- left_join(my_counties, Allison_rain, by = c("region", "subregion"))

rainmap %<>% mutate('Rain > 175 mm' = ifelse(rainmap$Rainfall < 175, "Unexposed", "Exposed"))
rainmap <- na.omit(rainmap)

ggplot() + 
  geom_polygon(data=my_counties, aes(x=long, y=lat, group=group),
               color="gray", fill="white", size = .1 ) + 
  geom_polygon(data = rainmap, aes(x = long, y = lat, group = group, 
                                   fill = `Rain > 175 mm`), 
               color = "grey", size = 0.2, alpha = 1.6) + 
  geom_polygon(data=my_states, aes(x=long, y=lat, group=group),
               color="black", fill="white",  size = 0.5, alpha = .3) + 
  geom_path(aes(x = Allison$longitude, y = Allison$latitude), color = "red", size=0.4) + 
  scale_fill_manual(values = c("darkblue", "white")) + 
  xlim(-107, -66) + 
  ylim(25, 48) + 
  ggtitle("Allison-2001") + 
  theme(plot.title = element_text(hjust = 0.5))
```


## The third map for Floyd-1999 made with tmap
```{r,warning=FALSE,fig.align='center'}
#fips match location
uscounty=st_as_sf(map('county',plot=F,fill=T))
colnames(county.fips)[2]=colnames(uscounty)[1]
uscounty=left_join(uscounty,county.fips,by='ID')

#floyd rain
rain_floyd=rain %>% filter(storm_id=='Floyd-1999')
total_rain_floyd = rain_floyd %>% group_by(fips) %>% summarise(storm_id=storm_id[1],precip=sum(precip)) 
total_rain_floyd = total_rain_floyd %>% mutate(fips=as.numeric(fips))
total_rain_floyd= right_join(uscounty,total_rain_floyd,'fips')

#floyd track line
hurr_floyd=hurr_tracks%>%filter(storm_id=="Floyd-1999")
track_floyd=cbind(longitude=hurr_floyd$longitude,latitude=hurr_floyd$latitude)
#the following codes refer to other's 
track_floyd=SpatialLines(list(Lines(Line(track_floyd),ID='Floyd-1999')))


#plot
tm_shape(total_rain_floyd)+
  tm_polygons('precip',title="Rainfall")+
  tm_layout(main.title="Floyd-1999 Rainfall",
            main.title.position="center") +
tm_shape(track_floyd)+
  tm_lines(col = "red",lwd = 3)
```

## The fourth map for Allison-2001 made with tmap

We extract data from rain dataset and match fips with list format of longitude and latitude. Besides, we extract data from hurr_tracks dataset and process data into format that can be read with tmap packages.

```{r,fig.align='center'}
## match fips with list format of longitude and latitude

data("county.fips")
ll<-st_as_sf(map("county", fill=T, plot = F))
county <- ll %>% 
  right_join(county.fips, by = c("ID" = "polyname"))

## extract data that storm id is Allison-2001, distance limit is 500 and rain limit is 175
allison_data<-filter_storm_data(storm="Allison-2001",distance_limit=500, rain_limit=175, include_rain = TRUE,days_included = -3:3, output_vars = c("fips","tot_precip")) 

allison_data$fips<-as.numeric(allison_data$fips)
allison_limit <- county %>%
  left_join(allison_data, by=c("fips"="fips"))
allison_limit<-na.omit(allison_limit)

## extract data that storm id is Allison-2001
allison_data_all<-filter_storm_data(storm="Allison-2001", include_rain = TRUE, days_included = -3:3, output_vars = c("fips","tot_precip")) 

allison_data_all$fips<-as.numeric(allison_data_all$fips)
allison_all <- county %>%
  right_join(allison_data_all, by=c("fips"="fips"))
allison_all<-na.omit(allison_all)
allison_all<-allison_all[1:2357,]

## hurricane track
data("hurr_tracks")
allison_hurr<-hurr_tracks %>%
  filter(storm_id=="Allison-2001")
allison_hurr_c=cbind(longitude=allison_hurr$longitude,latitude=allison_hurr$latitude)
allison_hurr_c<-SpatialLines(list(Lines(Line(allison_hurr_c),ID='Allison-2001')))

## map
tm_shape(allison_all)+
  tm_borders("black", lwd = .3)+
  tm_shape(allison_limit)+
  tm_polygons(col='tot_precip', title="rain>175mm")+
  tm_shape(allison_hurr_c) +
  tm_lines(col = "red",lwd = 1)+
  tm_layout(main.title="Allison-2001",
            main.title.position="center")
```






